climatrends: Precipitation and temperature indices for climate variability analysis in R

### 2020-05-20 inst/paper/paper.Rmd
fix-gdd-bug

\

@@

Four main functions are provided, `ETo()`, `GDD()`, `rainfall()` and `temperature()` with methods [@RCoreTeam] for classes 'data.frame', 'matrix' (or array), and 'sf' [@sf].

@@

@@

For cold environments, where $T_{min}$ may be lower than $T_{0}$, there are two variants of the previous equation to adjust either $T_{mean}$ (variant A) or $T_{min}$ (variant B). The variant A changes $T_{mean}$ to $T_{0}$ if $T_{mean} < T_{0}$ and is expressed as follows.

@@

The variant B, is calculated using Equation 1, but adjusts $T_{min}$ to $T_{0}$ if $T_{min} < T_{0}$, and is adjusted as follows.

@@

$$ T_{min} < T_{0} \; \rightarrow \; T_{min} = T_{0} $$

In `GDD()` these equations are defined as 'default', 'variant_a' and 'variant_b', respectively, and can be selected using the argument `equation`.

By default, the function returns the degree-days that is accumulated over the time series. Additionally, the function may return the daily values of degree-days or the number of days that a given organism required to reach a certain number of accumulated degree-days. These values are defined by 'acc', 'daily' or 'ndays' and can be adjusted using the argument `return.as`. The required accumulated gdd is defined with argument `degree.days`. For example, the Korean pine (*Pinus koraiensis*) requires 105 $^\circ C$ accumulated gdd to onset of photosynthesis [@JWu2013]. In that case, the function will calculate the growing degree-days ($gdd$) and sum up the values until it reaches the defined gdd (105 $^\circ C$) and return the number of days needed in the given season ($GDD_{ndays}$), as follows.

@@

@@

Where $p$ is the mean daily percentage of annual daytime hours, $T_{max}$ is the maximum temperature, $T_{min}$ is the minimum temperature, and $K_c$ is the factor for organism water need may change according to the growth stage.

>>>>

\
```{r get data, include=FALSE}
library("knitr")
library("kableExtra")
library("magrittr")
funtbl <- read.csv("available_functions.csv")
```

@@

Six main functions are provided, `crop_sensitive()`, `ETo()`, `GDD()`, `late_frost()`, `rainfall()` and `temperature()` with a default method for numeric 'vector's and additional methods [@RCoreTeam] for classes 'matrix' (or array), 'data.frame', and 'sf' (of geometry POINT or POLYGON) [@sf]. The later two are designed to fetch data from cloud sources.

@@

```{r characteristics, echo=FALSE, message=TRUE}
kable(funtbl, 
      row.names = FALSE,
      align = "l", 
      caption = "Main functions available in climatrends.") %>% 
  kable_styling(latex_options = "scale_down") %>% 
  column_spec(2, width = "1in") %>% 
  column_spec(3, width = "1.7in") %>% 
  column_spec(4, width = "1.7in")
```

@@

Additionally, the function `GDD()` offers three modified equations (two) designed for cold environments and (one) for tropical environments. For cold environments, where $T_{min}$ may be lower than $T_{0}$, there are two variants of the previous equation to adjust either $T_{mean}$ (variant a) or $T_{min}$ (variant b). The variant a changes $T_{mean}$ to $T_{0}$ if $T_{mean} < T_{0}$ and is expressed as follows.

@@

The variant b, is calculated using Equation 1, but adjusts $T_{min}$ or $T_{max}$ to $T_{0}$ if $T < T_{0}$, the equation is adjusted as follows.

@@

$$ T < T_{0} \; \rightarrow \; T = T_{0} $$

Where $T$ may refer to $T_{min}$ and/or $T_{max}$ when the condition of being below $T_{0}$ applies.

For tropical areas, where the temperature may surpass a maximum threshold ($T_{0^{max}}$), resulting in limited development, the minimum temperature is adjusted using Equation 3 and the maximum temperature is adjusted to a maximum base temperature as follow.

@@

$$ T_{max} > T_{0^{max}} \; \rightarrow \; T_{max} = T_{0^{max}} $$

Where $T_{0^{max}}$ is the maximum base temperature for growth, defined in `GDD()` using the argument `tbase_max`.

These modified equations are defined as 'a', 'b' and 'c', respectively, and can be selected using the argument `equation`.

By default, the function returns the degree-days that is accumulated over the time series using Equation 1. Additionally, the function may return the daily values of degree-days or the number of days that a given organism required to reach a certain number of accumulated degree-days. These values are defined by 'acc', 'daily' or 'ndays' and can be adjusted using the argument `return.as`. The required accumulated gdd is defined with argument `degree.days`. For example, the Korean pine (*Pinus koraiensis*) requires 105 $^\circ C$ accumulated gdd to onset the photosynthesis [@JWu2013]. In that case, `GDD()` will calculate the growing degree-days ($gdd$) and sum up the values until it reaches the defined gdd (105 $^\circ C$) and return the number of days needed in the given season ($GDD_{ndays}$), as follows.

@@

Where $p$ is the mean daily percentage of annual daytime hours, $T_{max}$ is the maximum temperature, $T_{min}$ is the minimum temperature, and $K_c$ is the factor for organism water need.
